Formalizing Stack Safety as a Security Property

### 2022-09-12 submission5/onlyformalism.tex
New model through IV

Security operations are split into two categories: {\it privilege operations}
and {\it control operations}. Privilege operations are those that adjust the
privileges of future callees by allocating private or public data and by saving
catch points for possible exceptions. Allocations specify the offset from
the stack pointer and size of the object. Exceptions are tied to the address
of the buffer that will contain saved state.
%
\begin{align*}
  \mathit{pop} \in \mathit{POPS} ::= & \mathbf{alloc} ~ \mathit{off} ~ sz & \mathit{off}, sz \in \mathbb{N} \\
  | & \mathbf{alloc\_pub} ~ \mathit{off} ~ sz & \mathit{off}, sz \in \mathbb{N} \\
  | & \mathbf{setex} ~ \addr_{buf} \\
\end{align*}

Control operations are those that exchange
control between principals: calls (including tailcalls), returns, and thrown
exceptions. Calls and tailcalls identify the target of the call, and the argument registers.
[Still a todo: stack arguments.] Return needs no further information. Throwing
an exception specifies the address of the saved buffer.
%
\begin{align*}
  \mathit{cop} \in \mathit{COPS} ::= & \mathbf{call} ~ \reg_{target} ~ \overline{\reg_{args}} & \\
  | & \mathbf{tailcall} ~ \reg_{target} ~ \overline{\reg_{args}} & \\
  | & \mathbf{return} & \\
  | & \mathbf{throw} ~ \addr_{buf} \\
\end{align*}
%
Security principals are function activations, which we model as an {\em activation record}.
Such a record will track the status of the principals data and the ways in which its callees
may return to it. To track the former, we define a set of {\it security classes}:
\[sc \in SEC ::= \sealed | \mathit{object} | \unsealed | \public\]

@@

We additionally define a set of {\it targets}, which identify how a
principal expects to have control returned to it. This could be via a normal
return to the correct instruction \(\addr_{pc}\), or by catching the exception
whose state is saved at \(\addr_{buf}\). The \(\mathit{Active}\) target indicates
the active principal.

\[\begin{split}
tar \in TAR ::= & \mathit{Return} ~ a_{pc} \\
| & \mathit{Catch} ~ a_{buf} ~ a_{pc} \\
| & \mathit{Active} \\
\end{split}\]

Finally, a {\it context stack} is a non-empty stack of triples of an activation,
a target, and a view. The initial context stack \(\context_0 = (\mathit{fresh},\mathit{Active},V_0)\),
and the top target on the stack will always be \(\mathit{Active}\).
\[\context \in \CONTEXTS ::= \mathit{list} ~ (A \times \mathit{TAR} \times \mathit{VIEW})\]

We define an update function for context stacks indexed by an overlay operation.
This will be defined for each operation as we introduce them in detail.
\[\constep{\psi} \in \MACHS \times \CONTEXTS \times \Psi \rightarrow \CONTEXTS\]

From a machine state and a context stack, we create a {\it combined state} \((\mach,\context)\)
and a transition that steps the context according to the label on the transition,
if there is one, and otherwise maintains it.

\judgmenttwo[Overlay]
            {\(\mach \harpoonunder{\psi} \mach'\)}
            {\(\mach',\context \constep{\psi} \context'\)}
            {\((\mach,\context) \stepstounder{\psi} (\mach',\context')\)}

\judgment[Default]
         {\(\mach \harpoonunder{\bot} \mach'\)}
         {\((\mach,\context) \stepstounder{} (\mach',\context)\)}

@@

In order to understand the motivation behind the call rule, though, we must first
describe how a function allocates objects. An allocation operation is parameterized
by a list of sizes; for each entry in the list, it marks an appropriate range
(based on offset from the stack pointer) as \(\mathit{object}\). This has no
direct bearing on the active principal, which may continue using the memory,
initialize it or not, and so on.

For calls, we define a function, {\it update view}, that defines the view
of a caller in terms of that of the callee. All argument registers are marked
\(\public\), along with all actual public registers. Callee-save registers,
including the stack pointer, are \(\sealed\); the callee is obligated to restore
them to their original value. The remaining caller-save registers are \(\unsealed\).

Then, all private objects are \(\sealed\), while the remainder of the stack retains
its old security context.

The formerly active principal is given a \(\mathit{Return}\) target, and on top of
it in the context stack we push a principal with a fresh identifier, the \(\mathit{Active}\)
target, and the updated view. On return, we search the stack for the first
\(\mathit{Return}\) target, and restore that principal to active status with its
original view. The rules given in Figure \ref{fig:callcontexts} take precedence
in order.

\begin{figure*}
  \judgment[AllocCNil]
           {\(\psi = \mathbf{alloc} ~ [ ~ ]\)}
           {\(\mach, \context \constep{\psi} \context\)}

  \judgmentthreebrtwo[AllocCCons]
                     {\(\psi = \mathbf{alloc} ~ s::sizes\)}
                     {\(\mach, (\alpha,T,V')::\context \constep{\psi'} \context'\)}
                     {\(b = \mach[sp] + sum(sizes)\)}
                     {\(\psi' = \mathbf{alloc} ~ sizes\)}
                     {\(V' = V\llbracket \addr \mapsto \mathit{object} | b \leq \addr < b+s \land V ~ \addr = \unsealed \rrbracket\)}
                     {\(\mach, (\alpha,T,V)::\context \constep{\psi} \context'\)}

  \[\mathit{reg\_view} ~ \overline{r_{args}} \triangleq \lambda \reg .
  \begin{cases}
    \public & \textnormal{if } \reg \in \overline{r_{args}} \textnormal{ or } \reg \in \mathit{PUBLIC} \\
    \unsealed & \textnormal{else if } \reg \in \mathit{CLR\_SV} \\
    \sealed & \textnormal{else} \\
  \end{cases}\]

@@

  \judgmenttwobr[CallC]
              {\(\psi = \mathbf{call} ~ r_{target} ~ \overline{\reg_{args}}\)}
              {\(T = \mathit{Return} ~ (\mach[pc] + 4)\)}
              {\(V' = \mathit{update\_view} ~ V ~ \overline{\reg_{args}}\)}
              {\(\alpha' = \mathit{fresh}\)}
              {\(\mach, (\alpha,\mathit{Active},V)::\context \constep{\psi}
                (\alpha',\mathit{Active},V')::(\alpha,T,V)::\context\)}

  \judgmenttwo[ReturnCRet]
              {\(\psi = \mathbf{return}\)}
              {\(T = \mathit{Return} ~ \_\)}
              {\(\mach, (\_, \mathit{Active}, \_)::(\alpha, T, V)::\context
                \constep{\psi} (\alpha, \mathit{Active}, V)::\context\)}

  \judgmenttwo[ReturnCSkip]
              {\(\psi = \mathbf{return}\)}
              {\(\mach, \context \constep{\psi} \context'\)}
              {\(\mach, (\alpha, \mathit{Active}, V)::\_::\context
                \constep{\psi} (\alpha, \mathit{Active}, V)::\context\)}

  \judgment[ReturnCNoRet]
           {\(\psi = \mathbf{return}\)}
           {\(\mach, [(\alpha,\mathit{Active},V)] \constep{\psi}
             [(\alpha,\mathit{Active},V)]\)}

\caption{Call and return context updates}
\label{fig:callcontexts}
\end{figure*}

@@

We convert macro-instructions into events via the \(\mathit{observe}\)
function, observing the values of relevant registers, defined in Figure \ref{fig:observe}.

@@

\begin{figure*}
\[\mathit{observe} ~ \psi? ~ m =
\begin{cases}
  \mathbf{callE} ~ m[\PCname] ~ m \llbracket r | r \in \overline{r_{args}} \rrbracket
  & \textnormal{when } ~ \psi = \mathbf{call} ~ r_{target} ~ \overline{r_{args}} \\
  \mathbf{callE} ~ m[\PCname] ~ m \llbracket r | r \in \overline{r_{args}} \rrbracket
  & \textnormal{when } ~ \psi = \mathbf{tailcall} ~ r_{target} ~ \overline{r_{args}} \\
%  (\rho, \mathbf{returnE} ~ m[\PCname] & \textnormal{when } \psi = \mathbf{return} \\
  \tau & \textnormal{otherwise} \\
\end{cases}\]
\caption{Converting overlay steps to events}
\label{fig:observe}
\end{figure*}

@@

\judgment{\(\not \exists \mach' ~ \context'. (\mach,\context) \stepstounder{\psi?} (\mach',\context')\)}
         {\((\mach,\context) \hookrightarrow \tau\)}

\judgmenttwo{\((\mach,\context) \stepstounder{\psi?} (\mach',\context')\)}
            {\((\mach',\context') \hookrightarrow \obsT\)}
            {\((\mach,\context) \hookrightarrow \notfinished{\mathit{observe} ~ \psi? ~ \mach}{\obsT}\)}

Naturally, most states will admit many possible traces, depending on the order in which
processors step.

We define another relation that takes a trace until we have returned to a particular
activation, or past it (as might occur when we add exceptions.)
We write this \(\context' \downarrow (\mach,\context) \hookrightarrow \obsT\), where \(\context'\)
is the context of the caller.

\judgment{\(|\context| \leq |\context'|\)}
         {\(\context' \downarrow (\mach,\context) \hookrightarrow \tau\)}

\judgmenttwobrlong{\((\mach,\context) \stepstounder{\psi?} (\mach',\context')\)}
                  {\(|\context'| > |\context''|\)}
                  {\(\context'' \downarrow (\mach',\context') \hookrightarrow \obsT\)}
                  {\(\context'' \downarrow (\mach,\context) \hookrightarrow \notfinished{\mathit{observe} ~ \psi? ~ \mach}{\obsT}\)}

@@

We start by defining a second-order logical operator
\(\context \uparrow P\), read ``\(P\) holds on return to \(\context\),''
where \(P\) is a predicate on machine states. This is a coinductive relation
similar to ``weak until'' in temporal logic -- it holds if the condition
is never reached.

\judgmenttwo[Returned]
            {\(|\context| \leq |\context_{ret}| }
            {\(P ~ (\mach,\context)\)}
            {\((\context_{ret} \uparrow P) ~ (\mach, \context)\)}

\judgmenttwobrlong[Step]
                  {\(|\context| \leq |\context_{tar}|\)}
                  {\((\context_{ret} \uparrow P) ~ (\mach', \context')\)}
                  {\((\mach, \context) \stepstounder{\psi?} (\mach', \context')\)}
                  {\((\context_{ret} \uparrow P) ~ (\mach, \context)\)}

Similarly, in {\it caller confidentiality}, we will want to compare future states,
so we give a binary equivalent, \(\context \Uparrow R\), so that
\((\mach,\context) ~ (\context_{ret} \Uparrow R) ~ (\mach',\context')\) holds if \(R\) holds on the
first return states after \((\mach,\context)\) and \((\mach',\context')\).

\judgmenttwobrlong[Returned]
            {\(|\context| \leq |\context_{ret}|\)}
            {\(|\context'| \leq |\context_{ret}|\)}
            {\((\mach,\context) ~ R ~ (\mach',\context')\)}
            {\((\mach,\context) ~ (\context_{ret} \Uparrow R) ~ (\mach',\context')\)}

\judgmenttwobrlong[Left]
              {\(|\context_1| \leq |\context_{ret}|\)}
              {\((\mach_1,\context_1) \stepstounder{\psi?} (\mach_1',\context_1')\)}
              {\((\mach_1',\context_1') ~ (\context_{ret} \Uparrow R) ~ (\mach_2,\context_2)\)}
              {\((\mach_1,\context_1) ~ (\context_{ret} \Uparrow R) ~ (\mach_2,\context_2)\)}

\judgmenttwobrlong[Right]
              {\(|\context_2| \leq |\context_{ret}|\)}
              {\((\mach_2,\context_2) \stepstounder{\psi?} (\mach_2',\context_2')\)}
              {\((\mach_1,\context_1) ~ (\context_{ret} \Uparrow R) ~ (\mach_2',\context_2')\)}
              {\((\mach_1,\context_1) ~ (\context_{ret} \Uparrow R) ~ (\mach_2,\context_2)\)}

>>>>

Security operations are split into two categories: {\it privilege operations}
and {\it control operations}. Privilege operations are those that adjust the
privileges of future callees by allocating private or public data and by saving
catch points for possible exceptions. Allocations specify the offset from
the stack pointer and size of the object. Exceptions are tied to the address
of the buffer that will contain saved state.
%
\begin{align*}
  \phi \in \Phi ::= & \mathbf{alloc} ~ \mathit{off ~ sz} & \mathit{off, sz} \in \mathbb{N} \\
  | & \mathbf{alloc\_pub} ~ \mathit{off ~ sz} & \mathit{off, sz} \in \mathbb{N} \\
  | & \mathbf{setex} ~ \addr_{buf} \\
\end{align*}

Control operations are those that exchange
control between principals: calls (including tailcalls), returns, and thrown
exceptions. Calls and tailcalls identify the target of the call, and the argument registers.
[Still a todo: stack arguments.] Return needs no further information. Throwing
an exception specifies the address of the saved buffer.
%
\begin{align*}
  \chi \in X ::= & \mathbf{call} ~ \reg_{target} ~ \overline{\reg_{args}} & \\
  | & \mathbf{tailcall} ~ \reg_{target} ~ \overline{\reg_{args}} & \\
  | & \mathbf{return} & \\
  | & \mathbf{throw} ~ \addr_{buf} \\
\end{align*}
%
A security operation may consist of any number of privilege operations and at most
one control operation.
%
\[\psi \in \Psi ::= \mathit{list ~ \Phi} | (X \times \mathit{list ~ \Phi})\]

Security principals are function activations, which we model as an {\em activation record}.
Such a record will track the status of the principal's data and the ways in which its callees
may return to it. To track the former, we define a set of {\it security classes}:
\[sc \in SEC ::= \sealed | \mathit{object} | \unsealed | \public\]

@@

An single principal might register multiple exception handlers over the
course of its execution, each of which will identify the value that the
program counter ought to have when an exception is thrown to it. We model
this as a partial map from addresses to addresses.
So an the activation record for the current activation is a pair of a view
and such a map:
\[(V,ex) \in \mathit{ACT} ::= \mathit{VIEW \times (\WORDS \rightharpoonup \WORDS)}\]

An inactive principal additional carries the address at which the program counter
should point in the event of a normal return.
\[(V, \addr_{ret}, ex) \in \mathit{INACT} ::= \mathit{VIEW \times \WORDS \times (\WORDS \rightharpoonup \WORDS)}\]

We represent a {\it call stack} as a pair of one active record and a list of inactive ones.
The initial call stack is \(\context_0 = ((V_0,[ ~ ]), [ ~ ])\).
\[(\alpha, \sigma) = \context \in \CONTEXTS ::= \mathit{ACT \times list ~ INACT}\]

Every security operation manipulates the call stack: privilege operations act on
the active principal, and control operations push or pop the active principal.
We define the security effects of each operation as a function \(Pop\) or \(Cop\),
from an operation, a machine state, and a call stack to a new call stack.
As we introduce more complex operations, we will separately define
the curried functions \(Pop ~ \phi\) or \(Cop ~ \chi\) for each operation.
They are combined in \(Op\), which acts on full operations as follows.
\[Pop \in \Phi \rightarrow \CONTEXTS \rightarrow \MACHS \rightarrow \CONTEXTS\]
\[Cop \in X \rightarrow \CONTEXTS \rightarrow \MACHS \rightarrow \CONTEXTS\]
\[Op ~ \psi =
\begin{cases}
  \lambda \context ~ \mach . \context & \textnormal{when } \psi = [~] \\
  Pop ~ \phi \circ Op ~ \overline{\phi} & \textnormal{when } \psi = \phi::\overline{\phi} \\
  Cop ~ \chi \circ Op ~ \overline{\phi} & \textnormal{when } \psi = (\chi,\overline{\phi}) \\
\end{cases}\]

A control operation may additionally produce an observable event, drawn
from a set \(\mathit{EVENTS}\) and ranged over by \(e\).
A function \(Ev\) maps operations and machine states to events. The
silent event \(\tau\) is relevant to our definitions here, and the
remainder will be described in Section \ref{sec:events}. 
%
\[Ev \in X \rightarrow \MACHS \rightarrow \mathit{EVENTS}\]
%
From a machine state and a context stack, we create a {\it combined state}
\((\mach,\context)\) and a transition \(\stepstounder{}\) on combined states,
labeled with an event. If the underlying machine transition is labeled by
some operation \(\psi\), the combined step applies \(Op\) to the contents
of \(\psi\) to determine the new call stack; otherwise the call stack is
unchanged. If there is a control operation,
the transition is labeled with its event, otherwise it is silent.
Rules apply in order.

\judgmentthree[Control]
              {\(\mach \harpoonunder{\psi} \mach'\)}
              {\(\psi = (\chi, \overline{\phi})\)}
              {\(e = Ev ~ \chi ~ \mach\)}
              {\((\mach,\context) \stepstounder{e} (\mach',Op ~ \psi ~ \mach ~ \context)\)}

\judgment[NonControl]
         {\(\mach \harpoonunder{\psi} \mach'\)}
         {\((\mach,\context) \stepstounder{\tau} (\mach',Op ~ \psi ~ \mach ~ \context)\)}

\judgment[Default]
         {\(\mach \harpoonunder{\bot} \mach'\)}
         {\((\mach,\context) \stepstounder{\tau} (\mach',\context)\)}

@@

In order to understand the motivation behind the call rule, though, we must first
describe how a function allocates objects. An allocation operation is parameterized
by an offset \(\mathit{off}\) and a size \(\mathit{sz}\). It will attempt to claim
a region of size \(\mathit{sz}\) starting at an offset of \(\mathit{off}\).
Claiming the region in this case means marking it as \(\mathit{object} ~ \mathbf{F}\),
where the boolean indicates whether the region is public.

\judgmenttwobrlong{\(b = \mach[sp] - \mathit{off}\)}
                  {\(\context = ((V, ex), \sigma)\)}
                  {\(V' = V \llbracket \addr \mapsto \mathit{object} ~ \mathbf{F} |
                    b \leq a < b+\mathit{sz} \land V ~ \addr = \unsealed \rrbracket\)}
                  {\(Pop ~ (\mathbf{alloc} ~ \mathit{off ~ sz}) ~ \context ~ \mach \triangleq
                    ((V',ex), \sigma)\)}

For calls, we define a function, {\it update view}, that defines the view
of a caller in terms of that of the callee. All argument registers are marked
\(\public\), along with all actual public registers. Callee-save registers,
including the stack pointer, are \(\sealed\); the callee is obligated to restore
them to their original value. The remaining caller-save registers are \(\unsealed\).
Then, all private objects are \(\sealed\), while the remainder of the stack retains
its old security context.

The formerly active principal's record is pushed onto the inactive list,
and replaced by a new record for the callee. Its return target is the intruction
immediately following that of the program counter before the call.
The callee's view is updated from the caller's with {\it update\_view}.

\judgmenttwo{\(\chi = \mathbf{call} ~ r_{target} ~ \overline{\reg_{args}}\)}
            {\(V' = \mathit{update\_view} ~ V ~ \overline{\reg_{args}}\)}
            {\(Cop ~ \chi ~ ((V,ex),\sigma) ~ \mach \triangleq
                ((V',[~]),(V, \mach[\PCname]+4, ex)::\sigma)\)}

On return, we restore the topmost inactive principal, if there is one.
Else, we default to retaining the current principal, as we do not model
the main function returning.

\judgmenttwo[Ret]
            {\(\chi = \mathbf{return}\)}
            {\(\sigma = (V,\addr_{ret},ex)::\sigma'\)}
            {\(Cop ~ \chi ~ (\_, \sigma) ~ \mach \triangleq
              ((V,ex), \sigma)\)}

\judgment[RetDef]
         {\(\chi = \mathbf{return}\)}
         {\(Cop ~ \chi ~ (\alpha, [~]) ~ \mach \triangleq
           (\alpha, [~])\)}

\begin{figure*}
  \[\mathit{reg\_view} ~ \overline{r_{args}} \triangleq \lambda \reg .
  \begin{cases}
    \public & \textnormal{if } \reg \in \overline{r_{args}} \textnormal{ or } \reg \in \mathit{PUBLIC} \\
    \unsealed & \textnormal{else if } \reg \in \mathit{CLR\_SV} \\
    \sealed & \textnormal{else} \\
  \end{cases}\]

@@

\caption{View updates for calls}
\label{fig:callcontexts}
\end{figure*}

@@

We convert control operations into events via the \(Ev\)
function, observing the values of relevant registers, defined in Figure \ref{fig:observe}.

@@

\begin{figure*}
\[\mathit{Ev} ~ \chi ~ m =
\begin{cases}
  \mathbf{callE} ~ (m[\PCname]+m[\reg_{target}]) ~ m \llbracket r | r \in \overline{r_{args}} \rrbracket
  & \textnormal{when } ~ \chi = \mathbf{call} ~ r_{target} ~ \overline{r_{args}} \\
  \mathbf{callE} ~ (m[\PCname]+m[\reg_{target}]) ~ m \llbracket r | r \in \overline{r_{args}} \rrbracket
  & \textnormal{when } ~ \chi = \mathbf{tailcall} ~ r_{target} ~ \overline{r_{args}} \\
  \tau & \textnormal{otherwise} \\
\end{cases}\]
\caption{Converting overlay steps to events}
\label{fig:observe}
\end{figure*}

@@

\judgmenttwo{\((\mach,\context) \stepstounder{e} (\mach',\context')\)}
            {\((\mach',\context') \hookrightarrow \obsT\)}
            {\((\mach,\context) \hookrightarrow \notfinished{e}{\obsT}\)}

We define another relation that takes a trace until we have returned to a particular
activation, or past it (as might occur when we add exceptions.)
We write this \(d \downarrow (\mach,\context) \hookrightarrow \obsT\), where
\(d\) is the depth of the original call.

\judgment{\(|\sigma| \leq d\)}
         {\(d \downarrow (\mach,(\alpha,\sigma)) \hookrightarrow \tau\)}

\judgmenttwobrlong{\((\mach,(\alpha,\sigma)) \stepstounder{e} (\mach',\context')\)}
                  {\(|\sigma| > d\)}
                  {\(d \downarrow (\mach',\context') \hookrightarrow \obsT\)}
                  {\(d \downarrow (\mach,(\alpha,\sigma)) \hookrightarrow \notfinished{e}{\obsT}\)}

@@

We start by defining a second-order logical operator
\(d \uparrow P\), read ``\(P\) holds on return to depth \(d\),''
where \(P\) is a predicate on machine states. This is a coinductive relation
similar to ``weak until'' in temporal logic -- it holds if the program never
returns to depth \(d\).

\judgmenttwo[Returned]
            {\(|\sigma| \leq d }
            {\(P ~ (\mach,(\alpha,\sigma))\)}
            {\((d \uparrow P) ~ (\mach, (\alpha,\sigma))\)}

\judgmenttwobrlong[Step]
                  {\(|\sigma| > d\)}
                  {\((d \uparrow P) ~ (\mach', \context')\)}
                  {\((\mach, (\alpha,\sigma)) \stepstounder{e} (\mach', \context')\)}
                  {\((d \uparrow P) ~ (\mach, (\alpha,\sigma))\)}

Similarly, in {\it caller confidentiality}, we will want to compare future states,
so we give a binary equivalent, \(d \Uparrow R\), so that
\((\mach,\context) ~ (d \Uparrow R) ~ (\mach',\context')\) holds if \(R\) holds on the
first states that return to depth \(d\) after \((\mach,\context)\) and \((\mach',\context')\).

\judgmenttwobrlong[Returned]
            {\(|\sigma_1| \leq d\)}
            {\(|\sigma_2| \leq d\)}
            {\((\mach_1,(\alpha_1,\sigma_1)) ~ R ~ (\mach_2,(\alpha_2,\sigma_2))\)}
            {\((\mach_1,(\alpha_1,\sigma_1)) ~ (d \Uparrow R) ~ (\mach_2,(\alpha_2,\sigma_2))\)}

\judgmenttwobrlong[Left]
              {\(|\sigma_1| > d\)}
              {\((\mach_1,(\alpha_1,\sigma_1)) \stepstounder{e} (\mach_1',\context_1')\)}
              {\((\mach_1',\context_1') ~ (d \Uparrow R) ~ (\mach_2,(\alpha_2,\sigma_2))\)}
              {\((\mach_1,(\alpha_1,\sigma_1)) ~ (d \Uparrow R) ~ (\mach_2,(\alpha_2,\sigma_2))\)}

\judgmenttwobrlong[Right]
              {\(|\sigma_2| > d\)}
              {\((\mach_2,(\alpha_2,\sigma_2)) \stepstounder{e} (\mach_2',\context_2')\)}
              {\((\mach_1,(\alpha_1,\sigma_1)) ~ (d \Uparrow R) ~ (\mach_2,\context_2')\)}
              {\((\mach_1,(\alpha_1,\sigma_1)) ~ (d \Uparrow R) ~ (\mach_2,(\alpha_2,\sigma_2))\)}
